---
title: "Function Friday: Regular Expressions"
output: html_document
author: "Thy Nguyen, Alex Denzler, & Hayley Hadges"
---
```{r, echo = FALSE}
library(stringr)
```

# Intro to Regular Expressions

Regular expressions are specially encoded text strings used as patterns for matching sets of strings, making them a concise and flexible tool for string-searching algorithms with "find" or "find and replace" operations on strings. 

Here are some real world regular expression applications:

* Searching files in a directory using the command line
* Checking for valid formatting (email address, phone number, date, etc)
* Control F 
* Parsing user input
* Replacing specific text
* Password pattern matching




Here is a [RegEx Cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) with a lot of this information condensed.


# Stringr Functions

While there are other popular R functions used for dealing with regular expressions, like `grep()`, `regexpr()`, and `gsub()`, we're going to focus on the functions in the `stringr` package. 

Function    |    Description
---------------- | ----------------------
`str_detect(string, pattern)`	| Detect the presence or absence of a pattern in a string, returns T/F.
`str_locate(string, pattern)`	| Locate the first position of a pattern and return a matrix with start and end.
`str_extract(string, pattern)` |	Extracts text corresponding to the first match, returns a vector.
`str_match(string, pattern)`	| Extracts capture groups formed by () from the first match.
`str_split(string, pattern)` |	Splits string into pieces and returns a list of character vectors.
`str_replace(string, pattern, replacement)`	 | Replaces the first matched pattern and returns a character vector.

# Basic Function Examples

#### **str_detect**
```{r}
month <- c("april", "April", "APRIL")
str_detect(month, "April")

months  <- c("March", "April", "May", "June", "July", "August")
str_detect(months, "a")
```

#### **str_locate()**
```{r}

```

#### **str_extract**
```{r}
x <- c("apple", "banana", "pear")
str_extract(x, "an")
```

#### **str_match()**
```{r}

```

#### **str_split()**
```{r}

```

#### **str_replace()**
```{r}

```



## Escape Characters

## Meta Characters

## Character Classes

## Quantifiers

## Anchors

As the default, regular expressions will match any part of a string. Anchors, on the other hand, do not match any character at all. Instead can anchor matches from the start or end of the given string.


`^` matches the start of the string.

`$`	matches the end of the string.


```{r}
x <- c("apple", "banana", "pear")
str_extract(x, "^a")

x <- c("apple pie", "apple", "apple cake")
str_extract(x, "apple$")

text = "the quick brown fox jumps over the lazy dog"
str_replace(text,"^the","---")
```




# Problem Set