---
title: "Function Friday: Regular Expressions"
output: html_document
author: "Thy Nguyen, Alex Denzler, & Hayley Hadges"
---
```{r, echo = FALSE}
library(stringr)
```

# Intro to Regular Expressions

Regular expressions are specially encoded text strings used as patterns for matching sets of strings, making them a concise and flexible tool for string-searching algorithms with "find" or "find and replace" operations on strings. Regular expressions may seem extremely complicated at first, however they can be used to save large amounts of time parsing through large text files. Regular expressions allow us to search for specific patterns within text, and we are even able to specify how many times we want to look for the pattern. Regular expressions are especially useful because they are able to be used across all mediums and coding languages.

Here are some real world regular expression applications:

* Searching files in a directory using the command line
* Checking for valid formatting (email address, phone number, date, etc)
* Parsing user input
* Replacing specific text
* Password pattern matching




Here is a [RegEx Cheatsheet](https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf) with a lot of this information condensed.

# Stringr Functions

While there are other popular R functions used for dealing with regular expressions, like `grep()`, `regexpr()`, and `gsub()`, we're going to focus on the functions in the `stringr` package. 

Function    |    Description
---------------- | ----------------------
`str_detect(string, pattern)`	| Detect the presence or absence of a pattern in a string, returns T/F.
*`str_locate(string, pattern)`	| Locate the first position of a pattern and return a matrix with start and end.
*`str_extract(string, pattern)` |	Extracts text corresponding to the first match, returns a vector.
*`str_match(string, pattern)`	| Extracts capture groups formed by () from the first match.
`str_split(string, pattern)` |	Splits string into pieces and returns a list of character vectors.
*`str_replace(string, pattern, replacement)`	 | Replaces the first matched pattern and returns a character vector.

Functions in the table with a `*` also have a `str_---_all()` variant that will match more than one occurrence of the pattern in a given string.

# Basic Function Examples

#### **str_detect**
```{r}
months  <- c("March", "April", "May", "June", "July", "August")
str_detect(months, "a")
```

#### **str_locate()**
```{r}
locations <- c("St Paul, MN, 55105", 
               "Minneapolis, Minnesota, 55404", 
               "Roseville, MN, 55113")

str_locate(locations, "MN")
```

#### **str_extract**
```{r}
fruits <- c("apple", "banana", "pear")
str_extract(fruits, "an")
```

#### **str_match()**
```{r}
str_match(locations, "MN")
```

#### **str_split()**
```{r}
str_split(locations, ",")
```

#### **str_replace()**
```{r}
str_replace(locations, "MN", "Minnesota")
```

## Meta Characters

Meta characters are the building blocks of regular expressions, as they allow for the regular expression to be written more concisely, specify character types, use ranges, and either specify or generalize what a regular expression is looking for. Below are examples of the commonly used meta characters.

Character        |    Description
---------------- | ----------------------
`*`	| Matches the preceding character 0 or more times. "`Be*`" matches "`B`", "`Be`", "`Bee`", "`Beee`", etc.
`+` | Matches the preceding character 1 or more times. "`Ze+`" matches "`Ze`", "`Zee`", "`Zeee`", etc.
`.` | Matches any alphanumeric or symbol. "`Da.`" matches "`Dab`", "`Da9`", "`Da@`", etc.
`?` | Matches the preceding character 1 or 0 times. "`To?n`" matches "`Ton`" or "`Tn`". 
`(...)` | Matches the characters or numbers in the exact order. "`Form(ula)`" matches "`Formula`" but not "`Formlau`". "`Form(123)`" matches "`Form123`" but not "`Form456`". 
`[...]` | Matches one of the characters or numbers. "`Alph[aeo]`" matches "`Alpha`" or "`Alpho`" or "`Alphe`". "`Alph[123]`" matches "`Alph1`", "`Alph2`", or "`Alph3`". This also works using "`-`" to show a range. "`Alph[a-z]`" allows any letter between "`a`" and "`z`" to follow the phrase "`Alph`". "`Alph[0-9]`" allows any number between "`0`" and "`9`" to follow the phrase "`Alph`". 
"`\W`" | Matches all symbols. "`\W`" matches "`$`", "`%`", "`#`", etc. "`\W\W`" matches "`#%`", "`&!`", etc.
"`\w`" | Matches all alphanumeric characters. "`\w`" matches "`B`", "`e`", "`4`", etc. "`\w\w\w`" matches "`bE1`", "`8Tn`", "`09s`", etc.
"`\d`" | Matches all whole numbers. "`\d`" matches "`1`", "`5`", "`0`", etc. "`\d\d\d\d`" matches "`1234`", "`8173`", "`0000`", etc. 

These can all be combined. For example, "`Tr.*`" matches "`Tr`", "`Tree`", "`Triumph`", "`Tragedy`", etc.

## Character Classes

A character class is a set of characters enclosed in the square brackets `[ ]`. A match is found when a character in the set is detected in the input. Some common character classes are:

Characters        |    Description
----------------- | ----------------------
`[abc]` | Matches any of the following characters `a`, `b`, or `c`. `gr[ae]y` matches `grey` and `gray`, and `[aeiou]` matches a lower case vowel while `[AEIOU]` matches an upper case vowel.
`[a-z]` | Matches a character in the alphabet (lower case), inclusive of `a` and `z`. 
`[0-9]` | Matches any digit from `0` to `9`, inclusive. `[4-9]th` matches `4th`, `5th`, `6th`, etc. but not `1th`, `2th` or `3th`.
`[a-zA-Z0-9]` | Matches any alphabetic character (lower or upper case) or single digit.
`[^...]` | Matches anything other than the specified characters. 

Character classes can be and are often used in conjunction with quantifiers and anchors (see below) to make regular expressions more flexible.

## Quantifiers

Quantifiers are the shortest to type, but they exercise immense power as they match the number of instances of a meta character, group, or character class in the input. If you want to repeat something in a regular expression, you can add a quantifier after it to specify how many times it can or should be repeated. The most common quantifiers are:

Quantifier        |    Description
---------------- | ----------------------
`*`	| Matches the preceding character 0 or more times. 
`+` | Matches the preceding character 1 or more times.
`?` | Matches the preceding character 1 or 0 times.
`{n}` | Matches the preceding character exactly n times.
`{n,}` | Matches the preceding character n or more times.
`{,n}` | Matches the preceding character at most n times.
`{n,m}` | Matches the preceding character at least n times but no more than m times.

The first three quantifiers are meta characters and are actually shorthands for the range quantifiers `{ }` listed below. The last quantifier, `{n,m}`, encompasses all of the above and understanding how it works is key to understanding the rest. This quantifier matches the preceding character (or group) n to m times. For example, a regular expression that can match the words "traveler" or "traveller" is "`travel{1,2}er`". `{,n}` is equivalent to `{0,n}` and `{n,}` is equivalent to `{n,Inf}`. When no comma is present as in the case of `{n}`, the pattern is matched exactly n times.

Thus, we can rewrite the first three quantifiers as:

Quantifier        | Normalized as range
---------------- | ----------------------
`*` | `{0,Inf}`
`+` | `{1,Inf}`
`?` | `{0,1}`

Some examples for the remaining quantifiers are:

Quantifier        |    Example
---------------- | ----------------------
`{n}` | "`Hel{2}o`" matches "`Hello`" but not "`Helo`", "`Heo`", "`Helllo`", etc.
`{n,}` | "`Sh{2,}`" matches "`Shh`", "`Shhh`", "`Shhhh`", etc. but not "`Sh`".
`{,n}` | "`\d{,4}`" matches numbers up to 4 digits and no numbers at all. Example matches are "" (empty string), "`8`", "`12`", "`694`", and "`2021`".
`{n,m}` | "`An{1,2}a`" matches only "`Ana`" and "`Anna`".

## Anchors

As the default, regular expressions will match any part of a string. Anchors, on the other hand, do not match any character at all. Instead, they match a position before or after given characters.


`^` matches the start of the string.

`$`	matches the end of the string.


```{r}
x <- c("apple", "banana", "pear", "apricot")
str_extract(x, "^a")
str_extract(x, "a$")

x <- c("apple pie", "apple", "apple cake")
str_extract(x, "apple$")
```


## Escape Characters

Because of the way that regular expressions are written, there is a specific way to look for characters such as `.`, `\`, `-`, and `?`, among others. There are specific ways to do this, using something called an escape character. An escape character tells the regular expression that you want to match the expression itself, not the behavior. In order to search for a literal `.`, your statement will look like this: `\.`. In this lies one issue. If you want to find a literal `\`, you must first escape it, creating the regular expression `\\`. In order to write the full expression, you must identify it is a string by opening and closing the expression using `\`. So to match `\`, your regular expression will look like this: `\\\\`. Here are some examples:

```{r}
x <- c("w.x.y.z", "wt[x")

str_detect(x, "wt\\[x")
str_detect(x, "w\\.x\\.y\\.z")
```

Here, we had to use an escape to identify the "`[`" and the "`.`", as they are outside of the character class.


# Problem Set

The dataset we will be using contains information on all New York Times bestselling hardcover fictions from 2009 through 2020. Please download the data using this [link](https://drive.google.com/file/d/12oqEt11miNGL_MtIcU75jx84pz6VEq_w/view?usp=sharing) and put it inside your working directory.

```{r, eval=FALSE}
# read in the data
bestsellers <- read.csv("bestsellers.csv") # the filepath to the downloaded data

bestsellers %>%
  head(10)
```


(@) Find books with multiple authors (HINT: Consider the possibility of an author having “and” in their name)

```{r, eval=FALSE}
str_???(bestsellers %>% pull(???), "???")
```


(@) Detect if the author’s first name starts with a vowel

```{r, eval=FALSE}
str_???(bestsellers %>% pull(author), "???")
```


(@) Change all authors with the name Alexander to Alex

```{r, eval=FALSE}
str_???_???(???, "Alexander", "Alex")
```

(@) Find books that are the second book in a series

```{r, eval=FALSE}
bestsellers %>%
  mutate(series = str_???(bestsellers %>% pull(???), pattern = "???")) %>%
  filter(series == TRUE)
```

(@) Find books that are the third or fourth one in a series

```{r, eval=FALSE}
bestsellers %>%
  mutate(series = str_???(bestsellers %>% pull(???), pattern = "???")) %>%
  filter(series == TRUE)
```

(@) Find books that are the 10th, 11th, ..., or 19th book in a series

```{r, eval=FALSE}
bestsellers %>%
  mutate(series = str_???(bestsellers %>% pull(???), pattern = "???")) %>%
  filter(series == TRUE)
```

(@) Describe in your own words how you would go about writing a regular expression for password pattern matching (ie 8 character minimum, one capital letter minimum, one lowercase letter minimum, one digit minimum, one special character minimum). This is an open ended question.








